The language of generalization: Probability, vagueness, and interaction

### 2018-09-24 paper/comgen-paper.Rmd
move prime numbered justices to gd

#### Worked examples

@@

\mht{this "surprising prediction" might set the reader up for the expectation that we are going to test this prediction..}

However, were we to learn much more surprising information---for instance, if every Supreme Court Justice in U.S. history had a social security number which was a *prime number* (a much lower probability outcome)---the shear suspiciousness could compel an observer to revise their theory of the domain (appealing perhaps to a conspiracy), update their subjective probability of future instances, and then endorse the generic. 
That is, our theory predicts there are situations where statements like (8) could be considered felicitous generalizations.

@@

### Cue validity

@@

The fact that conceptual knowledge is not a part of our formal semantics does not imply that such knowledge is unrelated to communicating generalizations.
Not only does conceptual knowledge influence interpretation via beliefs about prevalence, but conceptual knowledge may very well be what the speaker intends to communicate.
Like other models in the probabilistic pragmatic tradition, our model distinguishes the relevant variables for the truth-functional semantics from those that impact the speaker's utility in producing the utterance. 
A speaker's utility comes from addressing an implicit goal of communication, sometimes referred to as the *Question Under Discussion* or QUD [@roberts1996qud].
The separation of the QUD from the literal, truth-functional denotation of the utterance is an important theoretical distinction made explicit in our formal modeling approach [@Goodman2016]. 
For simplicity, the model presented here assumes the QUD is about prevalence.
However, the model makes nearly identical predictions if the QUD concerns the abstract parameters of the prevalence distributions [e.g., if the speaker intends to convey that "K is the kind of thing that Fs"; cf., @Prasada2006], something much closer to communicating conceptual understanding.
This observation may prove useful for constructing models of teachers, whose goals include conveying abstract relations between kinds and properties.

@@

Our view of understanding generalizations involving reasoning about an uncertain threshold is reminiscent of @Williamson1994's epistemicist view on vagueness.
This view holds that when a speaker utters a vague statement, there exists an objective set of criteria determining whether the entity in question satisfies the vague predicate.
Vagueness emerges from uncertainty about the details of those objective criteria (e.g., the precise value of a threshold). 
This uncertainty in turn could be the result of a population of speakers using slightly different criteria, thus shifting the criteria or threshold.
Listeners then will not know exactly what kind of speaker they are dealing with *a priori* and normatively should maintain uncertainty about the truth conditions. 
In our model, world knowledge and general communicative principles help a listener resolve the value of threshold.

@@

Perhaps the most surprising aspect of the language of generalization is how difficult it is to formalize, given how common it is in child-directed and child-produced speech [@Gelman1998; @GelmanEtAl2004].
<!-- from a formal perspective is that generic language is far from trivial to characterize.  -->
<!-- Generalizations are common in child-directed and child-produced speech [@Gelman1998; @GelmanEtAl2004] and are believed to be central to the growth of conceptual knowledge [@Gelman2004], -->
Generics are often contrasted with quantifier language (e.g., "some", "most", "all"), whose truth conditions are easy to formalize but which pose difficulty for young children to acquire [@Brandone2014; @Gelman2015].
<!-- If generics convey something about the probabilty of a feature, we would expect them to be in some way comparable to quantifier statments (e.g., "some", "most", "all", ...). -->
<!-- The extreme flexibility of generic meaning (e.g., "Birds lay eggs" vs. "Birds are female") stands in stark contrast to its early emergence in development. -->
<!-- In fact, quantified statements, whose formal meaning is much more straight-forward (e.g., "Some" means more than zero), emerge *later* in development [@Brandone2014; @Gelman2015], and are often confused for generic statements [@Leslie2012a]. -->
@Leslie2008 argues that the necessary complexity of a formal account of generics and the simplicity with which young children acquire generics implies that the normal tools for describing the semantics of quantified utterances (e.g., a truth-functional threshold) are inappropriate for generics.^[
  This argument is a little too simplistic in that it ignores evidence that children's struggle with quantifiers stems from pragmatic issues [@Musolino2006].
]

>>>>

### Worked examples

@@

@@

#### Cue validity

@@

The fact that conceptual knowledge is not a part of our formal semantics does not imply that such knowledge is unrelated to communicating generalizations.
In fact, we saw in Expt. 2c that it was participants predictions about the future that matter for the prevalence computation involved in endorsement.
Intuitive theories guide these predictions, and might provide new insight into classic puzzles of generics.
We described in the Worked Examples how our model treats "Supreme Court Justices have even social security numbers" as infelicitous even when 100% of justices have even SSNs, because a speaker's subjective probability that the *next justice* would have an even SSN is likely 50%.
However, were we to learn much more surprising information---for instance, if every Supreme Court Justice in U.S. history had a social security number which was a *prime number* (a much lower probability outcome)---the shear suspiciousness could compel an observer to revise their theory of the domain (appealing perhaps to a conspiracy), update their subjective probability of future instances, and then endorse the generic.

Not only does conceptual knowledge influence interpretation via beliefs about prevalence, but conceptual knowledge may very well be what the speaker intends to communicate.
Like other models in the probabilistic pragmatic tradition, our model distinguishes the relevant variables for the truth-functional semantics from those that impact the speaker's utility in producing the utterance. 
A speaker's utility comes from addressing an implicit goal of communication, sometimes referred to as the *Question Under Discussion* or QUD [@roberts1996qud].
The separation of the QUD from the literal, truth-functional denotation of the utterance is an important theoretical distinction made explicit in our formal modeling approach [@Goodman2016]. 
For simplicity, the model presented here assumes the QUD is about prevalence.
However, the model makes nearly identical predictions if the QUD concerns the abstract parameters of the prevalence distributions [e.g., if the speaker intends to convey that "K is the kind of thing that Fs"; cf., @Prasada2006], something much closer to communicating conceptual understanding.
This observation may prove useful for constructing models of teachers, whose goals include conveying abstract relations between kinds and properties.

@@

Our view of understanding generalizations involving reasoning about an uncertain threshold is reminiscent of @Williamson1992's epistemicist view on vagueness.
This view holds that when a speaker utters a vague statement, there exists an objective set of criteria determining whether the entity in question satisfies the vague predicate.
Vagueness emerges from uncertainty about the details of those objective criteria (e.g., the precise value of a threshold). 
This uncertainty in turn could be the result of a population of speakers using slightly different criteria, thus shifting the criteria or threshold.
Listeners then will not know exactly what kind of speaker they are dealing with *a priori* and normatively should maintain uncertainty about the truth conditions. 
In our model, world knowledge and general communicative principles help a listener resolve the value of threshold.

@@

Perhaps the most surprising aspect of the language of generalization is how difficult it is to formalize, given how common it is in child-directed and child-produced speech [@Gelman1998; @GelmanEtAl2004].
<!-- from a formal perspective is that generic language is far from trivial to characterize.  -->
<!-- Generalizations are common in child-directed and child-produced speech [@Gelman1998; @GelmanEtAl2004] and are believed to be central to the growth of conceptual knowledge [@Gelman2004], -->
Generics are often contrasted with quantifier language (e.g., "some", "most", "all"), whose truth conditions are easy to formalize but which pose difficulty for young children to acquire [@Brandone2014; @Gelman2015].
<!-- If generics convey something about the probabilty of a feature, we would expect them to be in some way comparable to quantifier statments (e.g., "some", "most", "all", ...). -->
<!-- The extreme flexibility of generic meaning (e.g., "Birds lay eggs" vs. "Birds are female") stands in stark contrast to its early emergence in development. -->
<!-- In fact, quantified statements, whose formal meaning is much more straight-forward (e.g., "Some" means more than zero), emerge *later* in development [@Brandone2014; @Gelman2015], and are often confused for generic statements [@Leslie2012a]. -->
@Leslie2008 argues that the necessary complexity of a formal account of generics and the simplicity with which young children acquire generics implies that the normal tools for describing the semantics of quantified utterances (e.g., a truth-functional threshold) are inappropriate for generics.^[
  This argument is simplistic in that it ignores evidence that children's struggle with quantifiers stems from pragmatic issues [@Musolino2006].
]
